---
title: "Before we Start"
teaching: 10
exercises: 5
questions:
- "FIX ME"
objectives:
- "FIX ME"
keypoints:
- "FIX ME"

source: Rmd
---

```{r, include = FALSE}
source("../bin/chunk-options.R")
knitr_fig_path("00-")
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Enron var et amerikansk energiselskab, der gik konkurs i 2001. De var med
22.000 ansatte en af verdens største energivirksomheder, og havde i 2000
en omsætning på 111 milliarder USD. Og seks år i træk betegnede finanstidsskriftet
Fortune Enron som det mest innovative selskab i USA. 

Lidt for innovative. Resulaterne skyldtes systematisk regnskabssvindel.

historien om datasættet.

## Taking a look at the data

We have the data in a zip-file, located in, if we followed the instructions in
the setup, the "data" folder in our project.

A zip file cannot be read directly. But R has a function called unzip() that will
unzip the file. We do that by typing unzip(). That is the function name. We 
also need to specify which file we want to unzip, including the path to the 
directory, so the unzip function knows where to find the file. 

The combined path and filename is what we call a string, or text. In R we 
indicate that something is text by placing it in quotation marks:

```{r eval=F}
unzip("data/emails.zip")
```

The unzip file will return the content of the zip-file. We know, because we have
prepared the zip-file in that way, that the content is a csv-file.

Looking at a spreadsheet, eg Excel, we see that we have rows, containing 
some observation, and columns, with variables

INDSÆT BILLEDE TIL ILLUSTRATION

The intersection between rows and colums are called cells.

Excel is a great tool, but it does not always play nice with other tools. A way
to get data out of excel is to export it as CSV, or Comma Separated Values.

A CSV-file is a text-file, only containing text, and with one line for each row
of the spreadsheet. Each cell is separated by a comma, hence Comma separated.

CSV files are easier to work with. But we still need to read it into R, so we 
can work with the data.

A good function for doing that is read_csv(), which, as the name implies takes
a csv-file as input, and reads it.

Now we need to get the output of the unzip function into the read_csv function.
There are several ways of doing this, but a nice way is using a pipe.

The pipe %>% takes what is on the left side of it, and sends it to what is on 
the right hand side of it.

The pipe is not native to R, but can be imported from external packages or
libraries. The concept is that someone smarter than us have written useful functions
that are not present in R as such, and distributed them in collections of functions,
or packages on the internet.

R have a robust way of importing packages, and the one we are going to work with
here is called tidyverse.

Getting to use at package comes in two steps. First we install it on our computer
using the function install.packages():
```{r eval = F}
install.packages("tidyverse")
```

Note that the name is a text-string, and in quotation marks.

Then we read in the package in R, so we can access it, using the function library():
```{r}
library(tidyverse)
```

Note that after we have installed tidyverse on our computer, tidyverse is no longer
a text string, but an object in R that we can call directly.

After doing this, we have access to the pipe (and read_csv btw), and can read in
the zip-file:

```{r}
df <- unzip("../data/emails.zip") %>% 
  read_csv()
```
We unzip the file, and use the pipe to send the output of that file to read_csv.
read_csv reads in the csv-file, and we store the data coming from the csv file
to an object in R, called df, using the assignment operator <-.

Think of that leftpointing arrow as, "put the result of what is on the right
hand side into what we write on the lefthand side.

df is an object, sometimes we call it a variable, which is kinda not entirely wrong.
We can store almost anything i objects, and we can call them almost everything we 
like.

So - what have we done so far?

We have read in more than ½ million emails from a zipped csv-file, to an object
in R, that we call df.

df is what we call a data frame, hence the abbreviation, and can be thought of
as the equivalent of a spreadsheet.

Let us take at look at the data. The function head() will show us the first six
rows:
```{r}
head(df)
```

We only see the first x columns, but there is a sixth, containing the actual content
of the email.

The columns contain the following data:

file - the original file that the mail was stored in. Interesting information 
might be stored here, because we can see that the first few rows are all from
the sent_mail folder. There is also deleted folders and others.
from - the e-mail of the person who sent the mail.
to - the e-mail(s) of the person(s) who the mail was sent to.
date - the date and time of the mail. In UTC. Most of the mails were actually sent from PDT
subject - the subject of the mail
content - the actual content of the mail.

But wait?! You wrote that this was big data and the original file is 1.4 Gb in size?

This file has been pared down. Basically all attachments have been removed.


